Results

### 2024-03-06 Analysis/discrim.qmd
cleanup; re-render; add some more reject model fits

---
title: "Testing Discrimination Analysis"
subtitle: "Fitting mixed effects models"
date: last-modified
categories: [Analysis, R]
code-fold: true
---

```{r}
#| output: false

lapply(c('tidyverse','data.table','lme4','lmerTest','knitr','kableExtra','cowplot','gghalves','here'),library,character.only = TRUE)
d <- readRDS(here("data/dPrune-07-27-23.rds"))

@@

```{r fig.width=11, fig.height=9}
#| eval: false
#| include: false
dtest %>% group_by(id,lowBound,condit,bandOrder,expMode) %>% summarise(n = n()) %>% ggplot(aes(x = n)) + geom_histogram(aes(fill = condit)) + facet_wrap(bandOrder+expMode~lowBound,ncol=6)

@@

## Quick Reminder of General Patterns

```{r fig.width=12, fig.height=10}
#| column: page-inset-right

fig1aCap=str_wrap("Figure 1a: Bands 100-300, 350-550 and 600-800 are novel extrapolations for both Original Order. Translucent rectangles indicate the correct band " ,width=170)

fig1bCap=str_wrap("Figure 1b: Bands 800-1000, 1000-1200,  and 1200-1400 are novel extrapolations for both Reverse Order. Translucent rectangles indicate the correct band " ,width=170)

plotDist <- function(df,title="",fcap=""){
  rectWidth=30
  df %>%ggplot()+aes(x = band, y = vxMeanCap, fill=vb) +
    # Set the color mapping in this layer so the points don't get a color
   geom_half_violin(color=NA)+ # remove border color
  geom_half_boxplot(position=position_nudge(x=-0.05),side="r",outlier.shape = NA,center=TRUE,
                    errorbar.draw = FALSE,width=20)+
  geom_half_point(transformation = position_jitter(width = 0.05, height = 0.05),size=.3,aes(color=vb))+
  facet_wrap(~condit,scale="free_x")+
    geom_rect(aes(xmin=band-rectWidth,xmax=band+rectWidth,ymin=band,ymax=highBound,fill=vb),alpha=.01)+
    geom_segment(aes(x=band-rectWidth,xend=band+rectWidth,y=highBound,yend=highBound),alpha=.8,linetype="dashed")+
    geom_segment(aes(x=band-rectWidth,xend=band+rectWidth,y=band,yend=band),alpha=.8,linetype="dashed")+
    labs(x = "Velocity Band", y = "vxMean",caption=fcap) +
    scale_y_continuous(expand=expansion(add=100),breaks=round(seq(0,2000,by=200),2))+
    scale_x_continuous(labels=sort(unique(df$band)),breaks=sort(unique(df$band)))+
    ggtitle(title) + theme(legend.position = "none")+theme_classic()+guides(fill="none",color="none")+
  theme(plot.caption=element_text(hjust=0,face="italic"))
}

#dtestAgg %>% plotDist()
dtestAgg %>% filter(bandOrder=="Original") %>% plotDist(title="Empirical Vx - Original Order",fcap=fig1aCap)
dtestAgg %>% filter(bandOrder=="Reverse") %>% plotDist(title="Empirical Vx - Reverse Order",fcap=fig1bCap)

@@

```{r fig.width=12, fig.height=9}
#| eval: false
#| include: false
#| 
#| column: page-inset-right
df_models <- bind_rows(df_pooled, df_no_pooling, df_partial_pooling) %>% 
  left_join(dtestAgg, by = c("id"))

# use base R to filter only testSlopeIndv with Rank < 3
testSlopeIndv[testSlopeIndv$rank<10,]$id

df_models %>% filter(id %in% testSlopeIndv[testSlopeIndv$rank<10,]$id) %>%
ggplot() + 
  aes(x = band, y = vxMed) + 
  # Set the color mapping in this layer so the points don't get a color
  geom_abline(
    aes(intercept = Intercept, slope = Slope_band, color = Model),
     linewidth= .75
  ) + 
  geom_point() +
  facet_wrap("id") +
  labs(x = "band", y = "vxMean") + 
  # Fix the color palette 
  scale_color_brewer(palette = "Dark2") + 
  theme(legend.position = "top", legend.justification = "left")
```

>>>>

---
title: "Testing Discrimination Analysis"
subtitle: "Fitting mixed effects models"
date: last-modified
#categories: [Analysis, R]
code-fold: true
---

```{r}
#| output: false
#| 
pacman::p_load('tidyverse','data.table','lme4','lmerTest','knitr','kableExtra','cowplot','gghalves','here')
d <- readRDS(here("data/dPrune-07-27-23.rds"))
source(here("Functions/Display_Functions.R"))

@@

```{r}
#| eval: false
#| include: false
dtest %>% group_by(id,lowBound,condit,bandOrder,expMode) %>% summarise(n = n()) %>% ggplot(aes(x = n)) + geom_histogram(aes(fill = condit)) + facet_wrap(bandOrder+expMode~lowBound,ncol=6)

@@

@@

```{r fig.width=12, fig.height=9}
#| column: page-inset-right
df_models <- bind_rows(df_pooled, df_no_pooling, df_partial_pooling) %>% 
  left_join(dtestAgg, by = c("id"))

#filter only testSlopeIndv with Rank < 3
testSlopeIndv[testSlopeIndv$rank<10,]$id

df_models %>% filter(id %in% testSlopeIndv[testSlopeIndv$rank<10,]$id) %>%
ggplot() + aes(x = band, y = vxMed) + 
  geom_abline(aes(intercept = Intercept, slope = Slope_band, color = Model),
     linewidth= .75) + 
  geom_point() + facet_wrap("id") +
  labs(x = "band", y = "vxMean")

### 2024-03-06 Analysis/e2_test.qmd
cleanup; re-render; add some more reject model fits

---
title: "HTW E2 Testing"
categories: [Analysis, R, Bayesian]
---

@@

result <- test_summary_table(testE2, "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kbl()
result$varied |> kbl()
# make kable table with smaller font size
# result$constant |> kbl(caption="Constant Testing - Deviation",booktabs=T,escape=F) |> kable_styling(font_size = 7)
```

@@

e2_distBMM |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600,800,1000,1200)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> gather_emmeans_draws()  |> 
   summarize(median_qi(.value),pd=sum(.value>0)/n()) |>
   select(contrast,Band=bandInt,value=y,lower=ymin,upper=ymax,pd) |> 
   mutate(across(where(is.numeric), \(x) round(x, 2)),
          pd=ifelse(value<0,1-pd,pd)) |>
   kbl(caption="Contrasts")

@@

result <- test_summary_table(testE2, "vx","X Velocity" ,mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kbl()
result$varied |> kbl()

>>>>

---
title: "HTW E2 Testing"
categories: [Analyses, R, Bayesian]
---

@@

result <- test_summary_table(testE2, "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kable()
result$varied |> kable()
# make kable table with smaller font size
# result$constant |> kbl(caption="Constant Testing - Deviation",booktabs=T,escape=F) |> kable_styling(font_size = 7)

@@

e2_distBMM |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600,800,1000,1200)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> gather_emmeans_draws()  |> 
   summarize(median_qi(.value),pd=sum(.value>0)/n()) |>
   select(contrast,Band=bandInt,value=y,lower=ymin,upper=ymax,pd) |> 
   mutate(across(where(is.numeric), \(x) round(x, 2)),
          pd=ifelse(value<0,1-pd,pd)) |>
   kable(caption="Contrasts")

@@

result <- test_summary_table(testE2, "vx","X Velocity" ,mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kable()
result$varied |> kable()

### 2024-03-06 Analysis/e3_test.qmd
cleanup; re-render; add some more reject model fits

---
title: "HTW E3 Testing"
categories: [Analysis, R, Bayesian]
---

@@

resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant 
resultOrig$varied

resultRev <- test_summary_table(testE3 |> filter(bandOrder=="Reverse"), "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
resultRev$constant 
resultRev$varied

@@

```{r}
#| label: tbl-e3-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant Testing - vx", "Varied Testing - vx"]
#| layout-ncol: 1
resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant 
resultOrig$varied

resultRev <- test_summary_table(testE3 |> filter(bandOrder=="Reverse"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultRev$constant 
resultRev$varied 
```

>>>>

---
title: "HTW E3 Testing"
categories: [Analyses, R, Bayesian]
date:  last-modified #"`r Sys.Date()`"
---

@@

resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant |> kable() 
resultOrig$varied |> kable()

resultRev <- test_summary_table(testE3 |> filter(bandOrder=="Reverse"), "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
resultRev$constant |> kable() 
resultRev$varied |> kable()

@@

```{r}
#| label: tbl-e3-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant Testing - vx", "Varied Testing - vx"]
#| layout-ncol: 1
resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant |> kable() 
resultOrig$varied |> kable()

resultRev <- test_summary_table(testE3 |> filter(bandOrder=="Reverse"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultRev$constant |> kable() 
resultRev$varied |> kable() 
```

### 2024-03-06 Model/htw_model.qmd
slight rearrange
rerender after reorganization; Results toc adjustments; remove multi-column and echo from indiv. results files.
cleanup; re-render; add some more reject model fits

---
title: HTW Model
author: Thomas Gorman
date: "`r Sys.Date()`"
page-layout: full
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

{{< include ../Discussion.qmd >}}

>>>>

---
title: HTW Model
author: Thomas Gorman
date:  last-modified #"`r Sys.Date()`"
page-layout: full
categories: [Modeling, ALM, EXAM, R]
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

{{< include ../Sections/Discussion.qmd >}}

### 2024-03-06 Model/htw_model_e2.qmd
cleanup; re-render; add some more reject model fits

---
title: HTW Model e2
author: Thomas Gorman
date: "`r Sys.Date()`"
page-layout: full
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

file_name <- "e2_n_iter_50_ntry_200_2506"
#file_name <- "n_iter_400_ntry_100_2944"
file_name <- "e2_n_iter_100_ntry_200_3436"

>>>>

---
title: HTW Model e2
author: Thomas Gorman
date:  last-modified #"`r Sys.Date()`"
page-layout: full
categories: [Modeling, ALM, EXAM, R]
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

#file_name <- "e2_n_iter_50_ntry_200_2506"
#file_name <- "n_iter_400_ntry_100_2944"
#file_name <- "e2_n_iter_100_ntry_200_3436"
file_name <- "e2_n_iter_200_ntry_300_2344"

### 2024-03-06 Model/htw_model_e3.qmd
cleanup; re-render; add some more reject model fits

---
title: HTW Model e3
author: Thomas Gorman
date: "`r Sys.Date()`"
page-layout: full
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

#file_name <- "e2_n_iter_50_ntry_200_2506"
#file_name <- "n_iter_400_ntry_100_2944"
#file_name <- "e2_n_iter_100_ntry_200_3436"
file_name <- "e3_n_iter_50_ntry_100_3244"

>>>>

---
title: HTW Model e3
author: Thomas Gorman
date:  last-modified #"`r Sys.Date()`"
page-layout: full
categories: [Modeling, ALM, EXAM, R]
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

#file_name <- "e2_n_iter_50_ntry_200_2506"
#file_name <- "n_iter_400_ntry_100_2944"
#file_name <- "e2_n_iter_100_ntry_200_3436"
#file_name <- "e3_n_iter_50_ntry_100_3244"
file_name <- "e3_n_iter_200_ntry_300_2033"

### 2024-03-06 Model/hybrid.qmd
cleanup; re-render; add some more reject model fits

---
title: HTW Hybrid Models
author: Thomas Gorman
date: "`r Sys.Date()`"
page-layout: full
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

ind_fits_df |> group_by(id,condit,Fit_Method) |> summarise(w=median(weight_exam),me=mean(mean_error))

ind_fits_df |> group_by(id,condit,Fit_Method) |> summarise(we=median(weight_exam),me=mean(mean_error)) |>
  group_by(condit,Fit_Method) |> summarise(w=mean(we),sd_w=sd(we),me=mean(me))

ind_fits_df |> 
  group_by(condit,Fit_Method) |> 
  summarise(w=median(weight_exam),sd_w=sd(weight_exam),me=mean(mean_error))

>>>>

---
title: HTW Hybrid Models
author: Thomas Gorman
date: "`r Sys.Date()`"
page-layout: full
categories: [Modeling, ALM, EXAM, R]
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

# ind_fits_df |> group_by(id,condit,Fit_Method) |> summarise(w=median(weight_exam),me=mean(mean_error))
# 
# ind_fits_df |> group_by(id,condit,Fit_Method) |> summarise(we=median(weight_exam),me=mean(mean_error)) |>
#   group_by(condit,Fit_Method) |> summarise(w=mean(we),sd_w=sd(we),me=mean(me)) 
# 
# ind_fits_df |> 
#   group_by(condit,Fit_Method) |> 
#   summarise(w=median(weight_exam),sd_w=sd(weight_exam),me=mean(mean_error))

### 2024-03-06 Sections/Intro.qmd
cleanup; re-render; add some more reject model fits

A longstanding issue across both science and instruction has been to understand how various aspects of an educational curriculum or training program influence learning acquisition and generalization. One such aspect, which has received a great deal of research attention,  is the variability of examples experienced during training.

Although varied training has often been purported to be particularly beneficial for generalization or transfer, few experiments have compared varied and constant training in contexts with unambiguous extrapolation testing.

@@

>>>>

A longstanding issue across both science and instruction has been to understand how various aspects of an educational curriculum or training program influence learning acquisition and generalization. One such aspect, which has received a great deal of research attention,  is the variability of examples experienced during training [@ravivHowVariabilityShapes2022]. The influence of training variation has been studied in numerous domains, including category learning [@cohenCategoryVariabilityExemplar2001; @posnerGenesisAbstractIdeas1968], visuomotor learning [@schmidtSchemaTheoryDiscrete1975; @bernikerEffectsTrainingBreadth2014 ],  language learning [@perryLearnLocallyThink2010], and education [@braithwaiteEffectsVariationPrior2015; @guoEffectsExampleVariability2014]. The pattern of results is complex, with numerous studies finding both beneficial  [@catalanoDistantTransferCoincident1984a; @braunMotorTaskVariation2009; @rollerVariablePracticeLenses2001], as well as null or negative effects [@huHighvariabilityTrainingDoes2024; @vanrossumSchmidtSchemaTheory1990; @brekelmansDoesHighVariability2022 ]. The present study seeks to contribute to the large body of existing research by examining the influence of variability in visuomotor function learning - a domain in which it has been relatively under-studied.

@@

### Variability and Function Learning

The influence of variability on function learning tasks has received relatively little attention.  @deloshExtrapolationSineQua1997 did include a variability manipulation (referred to as density in their paper), wherein participants were trained with either either 8, 20, or 50 unique input-output pairs (with the total number of training trials held constant). They found a minimal influence of variability on training performance, and no difference between groups in extrapolation.

@vandamMappingShapeVisuomotor2015 employed a visuomotor function learning task which required participants to learn a linear relationship between an input shape (which ranged continuously from very spiky to completely circular) and the correct pointing position.
